---
title: "Data Wrangling HW 9 and 10"
author: "Yaniv Bronshtein"
date: "4/19/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**
```{r echo=FALSE}
library(sna)
library(tidyverse)
library(igraph)
library(intergraph)
library(GGally)
library(rvest)
library(choroplethr)
library(Hmisc)
library(choroplethrMaps)
library(tidycensus)
library(gridExtra)
```

# Problem 1
1.Generate an undirected random graph with 20 vertices and probability of a tie
(forming an edge) equal 0.7. Calculate the mean degree from these graphs. 
Call it MD20.
2.Now, keeping the probability of a tie fixed at 0.7, repeat the above step with 
30, 40, 50, 60, 70, 80, 90 and 100 vertices, by writing a simple R function. 
In each case calculate the average mean degree. Denote them by MD30, ..., MD 100.
3.Plot the average mean degrees against the number of vertices. 
#Is the plot result consistent with your expectation? Explain.

**1.1 Generate MD20**
```{r}
set.seed(1)
net <- rgraph(n = 20, tprob = 0.7, mode = "graph")
MD20 <- mean(sna::degree(net, gmode = "graph"))
MD20
```
**1.2 repeat the above step with ** 
**30, 40, 50, 60, 70, 80, 90 and 100 vertices, by writing a simple R function.**
```{r}
get_mean_degree <- function(num_vertices) {
  set.seed(1)
  net <- rgraph(n = num_vertices, tprob = 0.7, mode = "graph")
  return(mean(sna::degree(net, gmode = "graph")))
}
```

**Call the function**
```{r}
mean_list <- NULL
for (i in seq(from = 20, to = 100, by = 10)) {
  mean_list <- c(mean_list, get_mean_degree(i))
}
```


**1.3 Plot the average mean degrees against the number of vertices.** 
**Is the plot result consistent with your expectation? Explain.** 
```{r}
#create the tibble
mean_degrees <- c("MD20","MD30", "MD40", "MD50", "MD60", "MD70", "MD80", "MD90", "MD100")
graph_means_t <- tibble(Vertices = seq(from = 20, to = 100, by = 10), 
                        MeanDegrees = mean_degrees, 
                        MeanValues = mean_list)
graph_means_t
```

**Create the plot**
```{r}
ggplot(data = graph_means_t, mapping = aes(x = Vertices, y = MeanValues)) +
  geom_point() +
  geom_line() + 
  labs(title = "Average mean degrees of Undirected graph vs number of vertices")


```

*The results indicate a linear relationship. As the number of of vertexes increase,*
*so does the average degree*



# Problem 2
1.Download the data frames "Dataset1-Media-Example-NODES.csv and 
"Dataset1-Media-Example-EDGES.csv" from 
https:// (Links to an external site.)net/network-visualization 
(Links to an external site.), and create an igraph network object.
2.Convert the igraph to a network object in the sna package
3. Plot the network using, using separate colors for the nodes based on the 
vertex attribute media.type and make the size of the nodes proportional to the 
vertex attribute audience.size. [Hint: Use network::get.vertex.attribute]
4. Calculate the mean degree and density of the network using appropriate 
functions in the sna package.


**2.1 Create an igraph from the data frame**
```{r}
nodes <- read_csv("C:\\Users\\Julia\\Downloads\\Dataset1-Media-Example-NODES.csv")
links <- read_csv("C:\\Users\\Julia\\Downloads\\Dataset1-Media-Example-EDGES.csv")
net_2 <- graph_from_data_frame(d=links, vertices = nodes, directed = TRUE) %>% 
  simplify(remove.multiple = F, remove.loops = T)

```

**2.2 Convert the igrpah to network object in the sna package**
```{r}
network_obj <- asNetwork(net_2)
```


**2.3 Plot the network using, using separate colors for the nodes based on the** 
**vertex attribute media.type and make the size of the nodes proportional to the** 
**vertex attribute audience.size. [Hint: Use network::get.vertex.attribute]**
```{r}
g_size <- network::get.vertex.attribute(network_obj, "audience.size")
g_color <- network::get.vertex.attribute(network_obj, "media.type")

ggnet2(network_obj, arrow.size = 5, arrow.gap = 0.02, size = g_size, 
       label = FALSE,  color = g_color, palette = "Set2")
```

**2.4 Calculate the mean degree and density of the network using appropriate** 
**functions in the sna package.**
**Density**
```{r}
density <- network.density(network_obj)
density
```
**Mean degree**
```{r}
mean_degree <- mean(sna::degree(network_obj, gmode = "digraph"))
mean_degree
```

# PROBLEM 3:
  
Scrape the country-wise population data from 
https://www.worldometers.info/world-population/population-by-country/ 
Plot the population density (P/Km2) obtained from this table on a 
country-wise choropleth map. Make sure to:

1.Clean the data to make it compatible with the country-wise world choropleth map
2.Maximize the overlap between the two data frames (the one obtained from the 
scraped data and the choropleth country data frame), 
i.e., if a country appears in both data frames, possibly with different names, it must be plotted
3.List the countries, if any, in the scraped data frame that do not appear in 
the choropleth country data frame (after appropriate cleaning)
4.List the countries, if any, in the choropleth country data frame that do 
not appear in the scraped data frame (after appropriate cleaning)

**Scrape data using rvest**
```{r}
data("country.regions") #Will be necessary as cross-reference
url <- "https://www.worldometers.info/world-population/population-by-country/"
worldometer_data <- url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table(fill = TRUE)

population_density_t <- worldometer_data[[1]] %>% dplyr::select(-1)
```

```{r}
clean_column <- function(data, col_name) {
  #Cast data as a tibble and extract the col_name as a vector
  col_vec <- as_tibble(data) %>% 
    pull(col_name)
  
  #If we are dealing with the value column, remove commas and convert to numeric
  if (col_name == "value") {
    return(
      col_vec %>%
        str_replace_all("[,]", "") %>%
        as.numeric())
  }
  #Else we are dealing with region column. We can only do initial processing here
  #which means lowercasing the column
  else {  
    return(
      col_vec %>% tolower()
    )
  }
} 
```

**Call the helper function to perform initial processing and select necessary columns**
```{r}

#Extract necessary region and value columns from population_density_t
cleaned_t <- population_density_t %>% 
  dplyr::select(1, 5)
names(cleaned_t)[1] <- "region"
names(cleaned_t)[2] <- "value"

#Call the clean_column() function to perform initial preprocessing on the data.
#Sort by the country/region ascending
cleaned_t <- cleaned_t %>% 
  mutate(region = clean_column(cleaned_t, "region"), 
         value = clean_column(cleaned_t, "value")) %>%
  arrange(region)
```

**use anti_join() to see the countries that are different between the country.regions**
**and the ones in our web scraped data**
```{r}
country_regions <- country.regions
misfits = anti_join(country_regions, cleaned_t)
misfits %>% head(10)


```

**Doing some outside work, we list the respective countries**
```{r}
choropleth_countries <- c("united states of america","republic of serbia",
               "united republic of tanzania","the bahamas",
               "democratic republic of the congo",
               "republic of congo","czech republic","guinea bissau",
               "ivory coast","macedonia","east timor","swaziland")
worldometer_countries <- c("united states","serbia","tanzania","bahamas","dr congo",
                 "congo","czech republic (czechia)","guinea-bissau",
                 "côte d'ivoire","north macedonia","timor-leste","eswatini")
```


**Replace the scraped versions of the misfits with the ones compatible with choropleth**
```{r}
for (i in seq(worldometer_countries)) {
  cleaned_t$region[cleaned_t$region==worldometer_countries[i]] <- 
    choropleth_countries[i]
}
```

**Below are the countries not present at all in the choropleth dataframe**
```{r}
misfits_choro <- anti_join(cleaned_t, country_regions)
misfits_choro %>% head(10)
nrow(misfits_choro)


```
**Generate the choropleth**
```{r}
choroplethr::country_choropleth(cleaned_t, title = "Population Density by Country")

```
# PROBLEM 4:
  
Obtain 2015-2019 5-year aggregated ACS tract-wise data on NJ median household 
income and rental. Combine them into a single data frame.

1.Plot the tract-wise rental against the median household income and comment.
2.Fit a linear regression equation of rental against median household income 
and report the summary.
3.Looking at the plot, suggest ways to improve the model fit. 
Fit the improved model and report the R2.
4.Obtain the rental data for year=2019 (5-year aggregate from 2015-2019) 
and year=2014 (5-year aggregate from 2010-2014), and plot the percentage 
changes for each county in a column or bar diagram, in increasing or
decreasing order of percentage increase.

**4.1 Plot the tract-wise rental against median household income**


```{r}
my_census_key <- "68b91a0cb82e99d93fcf934623a6b9ad5ba04986"
census_api_key(my_census_key, install = TRUE, overwrite = TRUE)

v19 <- load_variables(2019, "acs5", cache = TRUE) #Load the variables

acs_data_t <- get_acs(geography = "tract",
                      variables = c(medincome = "B19013_001", 
                                    medrental = "B25064_001"),
                      state = "NJ")

acs_data_t <- acs_data_t %>% select(NAME, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate)

acs_data_t <- acs_data_t %>% filter(!is.na(medrental), !is.na(medincome))
ggplot(data = acs_data_t, mapping = aes(x = medincome, y = medrental)) +
  geom_point() + 
  geom_smooth(method = "lm")
```
*It seems as though the two variables are positively correlated*


**4.2Fit a linear regression equation of rental against median household income** 
```{r}
acs_lm <- lm(medrental ~ medincome, data = acs_data_t)

summary_lm <- summary(acs_lm)
lm_rsq <- summary_lm$r.squared
lm_rsq
```

**4.3 Looking at the plot, suggest ways to improve the model fit.**
**Looking at the plot, suggest ways to improve the model fit.**
```{r}
acs_poly <- lm(medrental ~ poly(medincome, 5), data = acs_data_t)
summary_poly <- summary(acs_poly)
poly_rsq <- summary_poly$r.squared
poly_rsq
```
*Fitting a poly model with order=5 increased the R^2 by almost 0.01*


**4.4 Obtain the rental data for year=2019 (5-year aggregate from 2015-2019)** 
**and year=2014 (5-year aggregate from 2010-2014), and plot the percentage** 
**changes for each county in a column or bar diagram, in increasing or**
**decreasing order of percentage increase.**

**Get the 2014 and 2019 data**
```{r}
rental_14 <- get_acs(geography = "county",
                     variables = c(medrental= "B25064_001"), 
                     year = 2014, state = "NJ") %>% 
  rename(region = NAME, value = estimate)


rental_19 <- get_acs(geography = "county",
                     variables = c(medrental= "B25064_001"), 
                     year = 2019, state = "NJ") %>% 
  rename(region = NAME, value = estimate)
```

**Use choroplethr to calculate the percent change and store in a dataframe**
```{r}
percent_change_t <- choroplethr::calculate_percent_change(rental_14, rental_19)
```


**Create a ggplot() column graph to show to show information on rental changes**
```{r}
ggplot(data = percent_change_t,
       mapping = aes(x = value, y = fct_reorder(region,value))) + 
  geom_col(fill = "magenta") +
  labs(y="County",
       x="% Change in rent",
       title = "Percentage change of rentals in NJ 2014-2019")
```
