---
title: "Assignment 2 Data Wrangling"
author: "Yaniv Bronshtein"
date: "02/07/2021"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the libraries
```{r}
library(tidyverse)
library(gapminder)
```
# 1. Download the dataset on restaurant inspection in csv format from
# https://data.cityofnewyork.us/Health/DOHMH-New-York-City-Restaurant-Inspection-Results/43nn-# pn8j (Links to an external site.)
# and convert it to a data frame.
**read data from csv remove leading and trailing White space**
```{r}
res_inspec_results <- read_csv(
  "DOHMH_New_York_City_Restaurant_Inspection_Results.csv",
  trim_ws = TRUE)
```

**Convert data read from file to data frame**
```{r}
df <- as.data.frame(res_inspec_results)
```


## (1a) Form a new data frame restricted to restaurants in Queens with cuisine 
## to "Pizza". [Hint: Use str_detect() in the library stringr. 
## Examples of usage of str_detect() can be found here: 
## https://stringr.tidyverse.org/reference/str_detect.html]

  
**Filter df by specifying the BORO field as Queens and using str_detect()**
**applied in conjunction with fixed() to ascertain that only the rows in the**
**original data frame relating to Pizza in the borough Queens are left are** 
**left**
```{r}
queens_pizza <- filter(df, BORO == "Queens", 
                         str_detect(`CUISINE DESCRIPTION`,
                                    fixed("Pizza", ignore_case = TRUE))) 
```


## 1b). What are the 5 most commonly names inspected restaurants 
## (use the variable "DBA") in the data frame? (queens_pizza)

** Perform the following steps**
**1. Group the queens_pizza data frame by the DBA field**
**2. Pipe the result of step 1 to summarize which will return a new data frame** 
**with just the frequency and the DBA field**
**3. Pipe the result of step 2 to the ungroup() function**
**4. Pipe the result of step 3 to the arrange() function to sort the frequency**
**in descending order**
**5. Pipe the result of step 4 to the head() function which is passed the**
**constant 5 to display only the first 5 rows of the data frame**
**6. Store the result of steps 1 through 5 in the variable common_names**
```{r}
common_names <- queens_pizza %>% group_by(DBA) %>% summarise(n=n()) %>% 
  ungroup() %>% arrange(-n)%>%head(5)

```
**Plot the results in a bar graph**
```{r}
ggplot(data = common_names,
       mapping = aes(x = DBA, y = n)) +
  geom_bar(stat = 'identity', color = "green", fill="purple") +
  labs(
    title = "5 most commonly inspected restaurant names",
    x = "Restaurant Name",
    y = "Count"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

*Based on the data in the graph, the 5 most inspected restaurants are DOMINOS'*
*PAPA JOHN'S, Rosa's Pizza, PIZZA HUT, AND 3 BROS 99 CENT PIZZA* 

##1c). On what dates has queens pizza parlor "SUSANO'S PIZZERIA & RESTAURANT"
## been inspected?
**Filter queens_pizza to match the exact restaurant name. NOTE:**
**we cannot use str_detect() because that would mean including restaurants**
**such as SUSANO'S PIZZERIA & RESTAURANT I**

```{r}
queens_pizza_susanos <- filter(queens_pizza,DBA=="SUSANO'S PIZZERIA & RESTAURANT")
```

**Next, we make sure to call distinct() to get rid of duplicate dates. We**
**only need the INSPECTION DATE column**
```{r}
dates_inspected <- distinct(queens_pizza_susanos["INSPECTION DATE"])
dates_inspected
```

# 2. The file "gapminder_2007_gini.tsv" is in the “Assignments” folder under the files menu of # the course website; it is a subset of the 2007 Gapminder data merged with recent Gini
# coefficient data (https://en.wikipedia.org/wiki/Gini_coefficient

## 2a). Create a plot to compare the distributions (e.g., central tendency, dispersion) of the ## Gini coefficient in different continents. [Hint: Use using geom_boxplot(). Details and 
## examples are available here: https://ggplot2.tidyverse.org/reference/geom_boxplot.html]


**Read in the data from the tab separated vector file, trimming white space**
```{r}
gapminder_data <- read_tsv("gapminder_2007_gini.tsv", trim_ws = TRUE)
```
**Convert the data read from the file to a data frame**
```{r}
gap_df <- as.data.frame(gapminder_data)
```

**Create the box plot of distribution by continent**
```{r}
ggplot(data = gap_df, mapping = aes(x = continent, y = gini)) +
  geom_boxplot() +
  labs(
    title = "Gini Distribution by Continent"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```

## 2b) Does the Gini coefficient appear to have any impact on the life 
## expectancy in 2007? Explain your answer using a plot, 
## classified by continents
**1. Filter the gap_df data frame by the year 2007**
**2. Pipe the result of step 1 to group_by() to aggregate the data by continent** 
**3. Pipe the result of step 2 to the summarise() function which returns**
**a data frame with 3 columns: the continents, the median of the gini, and the**
**median of the life expectancy**
**4. Store the result of steps 1 through 3 in the variable gini_vs_life_exp**

```{r}
gini_vs_life_exp <- gap_df %>% filter(year == 2007)  %>% group_by(continent)  %>%
  summarise(continent, gini, lifeExp)
```

**Create a bar plot using the gini_vs_life_exp data frame**
```{r}

ggplot(data = gini_vs_life_exp, mapping = aes(x = gini, y = lifeExp, color = continent)) +
  geom_point() +
  geom_smooth(method = "lm", mapping = aes(group = 1)) +
  
  labs(
    title = "Effect of gini on life expectancy by Continent"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```
*The graph above demonstrates a trend that countries with low gini such as*
*Europe and Oceania have much higher life expectancy than Africa which falls*
*below the linear regression line*

# (3) Using the original gapminder data frame, please generate a data frame 
# with a new variable called gdp by multiplying the population size by the gdp 
# per capita. To make those large numbers more understandable, please form an 
# new variable called gdp_ratio equal to the gdp divided by the gdp of the 
# United States in 2007.
# Find the mean gdp_ratio by continent and year, and then plot the mean 
# gdp_ratio over time, distinguishing the continents. 
# Please use both points and lines for the plot

**Create a data frame to store the imported data from the R package gapminder**
```{r}
df2 <- gapminder
```

**Call the dplyr function mutate() to create a calculated column in df2**
**to store the gdp**
```{r}
df2 <- dplyr::mutate(df2, gdp = pop * gdpPercap)
```

**To avoid writing extra code, extract the 2007 gap data on the United States**  
```{r}
us_2007_df <- gap_df %>% filter(country=="United States")

```

**Calculate the gdp manually from the single row by multiply the gdp per capita**
**by the population at the time**
```{r}
us_gdp_2007 <- us_2007_df["gdpPercap"] * us_2007_df["pop"]  
```
**Convert the result to a double value and store in the variable denom**
```{r}
denom <- as.numeric(us_gdp_2007)
```
**Create another calculated column in df2 by calling mutate() on the calculation**
**gdp / denom**
```{r}
df2 <- dplyr::mutate(df2, gdp_ratio = gdp / denom )
```

**1. Pipe df2 to the function group_by() to group by continent and year**
**2. Pipe the result of step 1 to the summarise() function which returns**
**a data frame with 3 columns: the continents, the year the mean of the gdp ratio**
**3. Save the results in the variable q3_info
```{r}
q3_info <- df2 %>% group_by(continent, year)  %>%
  summarise(continent, year, mean_gdp_ratios = mean(gdp_ratio))
```

**Create the plot of all the data**
```{r}
ggplot(data = q3_info, mapping = aes(x = year, y = mean_gdp_ratios, color = continent)) +
  geom_point() +
  geom_line()
```






