---
title: "Second Markdown Example"
author: "Tirthankar Dasgupta"
date: "January 25, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
```

## Some Notes on Flights

In this document, we will:

* Try some markdown code, including 
    + headers
    + items
* Include some inline R code
* Include a plot

The flights data.frame in the nycflights13 package has data on `r nrow(flights)` flights. Some of the flights have missing data, denoted by "NA". That means if we type
```{r}
mean(flights$dep_delay)
```
the answer is NA, as shown above.

Instead, type
```{r}
mean(flights$dep_delay, na.rm = TRUE)
```
(The mean function has an option called na.rm that, when set to TRUE, will remove NAs and then calculate the mean.) 

The mean flight delay is `r mean(flights$dep_delay, na.rm = TRUE)`. The distribution of delays can be visualized with a histogram, shown below,



```{r chunk1, warning = FALSE}
ggplot(data = flights) + geom_histogram(mapping = aes(dep_delay))

```




or as a density plot.

```{r density}
ggplot(data = flights) + geom_density(mapping = aes(dep_delay)) 
```


It's important to note that your R code either goes inside the triple ticks, \`\`\`, or the single tick, \`, and don't forget the "r".

