---
title: "Assignment 1 Data Wrangling"
author: "Yaniv Bronshtein"
date: "1/31/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the libraries
```{r}
library(babynames)
library(tidyverse)
```
## Plot the number of male and female babies named Taylor *by year*
**Filter the babynames dataset for the name Taylor**
```{r, echo=FALSE}
baby_Taylor = filter(babynames, name=="Taylor")
```
**Create a ggplot using the newly filtered data where the x-axis is the year,**
**y-axis is the count of the number of babies, and the color of each line is** 
**defined by the sex variable**


```{r, echo=FALSE}
ggplot(data = baby_Taylor) +
  geom_line(mapping=aes(x=year, y=n, color=sex)) +
  labs(title = "Male and Female babies named Taylor by year",
       x = "years",
       y = "# of babies")
```


# Answer the following questions, showing plots to substantiate your answers: 
 
## 1. Is a 16 year old named Quinn more likely to be a boy or a girl?
**Initialize the variables for the scale**
```{r}
max_year <- max(babynames$year)
year_16 <- max_year - 16
```
**Filter the babynames data set for the name Quinn, and the year is 2001 onward**
```{r}
baby_Quinn = filter(babynames, name=="Quinn", year>= year_16)
```
**Create a ggplot using the newly filtered data where the x-axis is the year,**
**y-axis is the count of the number of babies, and the color of each line is** 
**defined by the sex variable. The scale is set at 2 year increments starting** 
**2001 and ending 2017**
```{r}
ggplot(data = baby_Quinn) +
  geom_line(mapping=aes(x=year, y=n, color=sex)) +
  labs(title = "Male vs Female Trend of baby name Quinn",
       x = "Year",
       y = "# of babies") +
  scale_x_continuous(breaks = seq(year_16, max_year, by = 2))
```

*According to the plot above, a 16 year old named Quinn was born in 2001 since* 
*the data set ends in 2017. Even if one were to consider the current year to be*
*2021(which would mean Quinn was born in 2005,Quinn was still mostly a boy’s name.)*

### 2. Is a 2 year old named Quinn more likely to be a boy or a girl?
*According to the plot above, a 2 year named Quinn would have been born in 2015.* 
*During this time, Quinn was overwhelmingly a girl’s name. Since the data ends*
*in 2017, it is impossible to consider the current year to be 2021 with a baby* 
*born in 2019.*

### 3. What is your best guess as to how old a woman named Susan is?
**Filter the babynames dataset for the name Taylor**
```{r}
susan_female <- filter(babynames, name=="Susan", sex == "F")
```
**Create a ggplot using the newly filtered data where the x-axis is the year,**
**y-axis is the count of the number of babies**
```{r}
ggplot(data = susan_female) +
  geom_line(mapping=aes(x=year, y=n)) +
  labs(title = "Trend for a female for the baby name Susan",
       x = "Year",
       y = "# of babies")
```

*According to the plot above, the popularity of the baby name Susan can be* 
*approximated by the normal distribution with a peak popularity between 1950*
*and 1960. If one takes 2017 to be the final year, that would put our Susan* 
*at between 57 and 67.  I believe the question to be ambiguous as even in 1980*
*there are roughly 5000 Susans born. For all we know, there was an attack on*
*the Susans born between 1950 and 1960 and the only Susans left were born in*
*1980 making them 37 years old.*

