---
title: "Assignment4"
author: "Yaniv Bronshtein"
date: "2/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Import the libraries
```{r}
library(tidyverse)
```

## Question 1

Create a new function that, given an `lm` object, returns the top n residuals  
arranged in descending order according to their largest absolute values  
(but returns the residuals, not the absolute value of the residuals),  
where the default value for n is 5. The function should give a clear error  
message if n is larger than the number of residuals.   
Demonstrate that your function works by applying it to
mtcars.lm <- lm(mpg ~ disp, data = mtcars)
first with no argument for n, then with n = 6, and then with n = 40 (error message expected).

```{r cars}
sort_res_by_abs <- function(obj, n = 5) {
  if (!("lm" %in% class(obj))) stop ("obj must have class 'lm'")
  num_res <- length(obj$residuals)
  res <- as_tibble(obj$residuals)
  print("Num res")
  print(num_res)
  if (n > num_res) {
    stop (paste("the n value", n, "is larger than the number of residuals", 
                num_res, sep = " "))
  } 
  sorted_res <- res %>% rename(res = value) %>% 
    mutate(abs_res = abs(res)) %>%
    arrange(desc(abs_res)) %>% 
    select(res) %>% head(n)
  return(as.list(sorted_res))
  
}

mtcars.lm <- lm(mpg ~ disp, data = mtcars)
sort_res_by_abs(mtcars.lm, 20)
```

## Question 2
Read the file "height.txt" in the folder, and use regular expressions to clean the height variable and convert it into a numeric column representing height in inches. Determine and report the number of non-missing values of height for men and women. Finally, make a plot showing the distributions of height for men and for women on the same plot.

