---
title: "Assignment 7 Data Wrangling"
author: "Yaniv Bronshtein"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**
```{r}
library(tidyverse)
library(rvest)
library(broom)
library(Hmisc)
library(ggrepel)
library(jsonlite)
```
# PROBLEM 1:
**1.From the worldometer webpage** 
**https://www.worldometers.info/coronavirus/usa/new-jersey/**
**extract the county-wise COVID data (total cases, new cases, total deaths and new deaths).** 
**2.Show a nice graphical representation of the county-wise total cases and** 
**total deaths in a single plot. Use your imagination and Chapter 3 of R for** 
**Data Science to come up with an appropriate visual representation.**
**3.Identify the top two counties reporting most new cases.** 


**Part 1**
```{r}
url <- "https://www.worldometers.info/coronavirus/usa/new-york/"

worldometer_t <- url %>%
  read_html() %>%
  html_nodes("table") %>%
  html_table(fill = TRUE)
```

**Use yesterday's coronavirus data for New York State since it is fully updated**
**Perform the following steps::**
**a). Convert worldometer_t[[2]] which is yesterdays data to a tibble**
**b). From this tibble extract only the County, TotalCases, NewCases, TotalDeaths, and NewDeaths**
**c). Remove the rows that contain the string Total signifying state totals since we only**
**want county-wide data**
```{r}
county_covid_t <- worldometer_t[[2]] %>% 
  as_tibble() %>%
  select(County, TotalCases, NewCases, TotalDeaths, NewDeaths) %>%
  filter(!str_detect(County ,"Total"))
```

**Create a function clean_column that takes in the tibble data and the column to**
**clean and returns a numeric vector of the cleaned up column**
```{r}
clean_column <- function(data, col_name) {
  #Cast data as a tibble and extract the col_name as a vector
  col_vec <- as_tibble(data) %>% 
    pull(col_name)
  
  #Check if the vector is numeric. If not, the data requires processing
  #All commas and + need to be removed, the vector needs to be converted to numeric,
  #and the NA's need to be replaced with 0's
  if (!all.is.numeric(col_vec)) {
    return(
      col_vec %>%
        str_replace_all("[,+]", "") %>%
        as.numeric() %>%
        replace_na(0)
    )
  } 
  #If the data is already numeric, only the NA's need to be replaced with 0's
  else {
    return(
      col_vec %>%
        replace_na(0)
    )
  }
}
```

**Call clean_column() on each column of county_covid_t**
```{r}
total_cases <- clean_column(data = county_covid_t, col_name = "TotalCases")
new_cases <- clean_column(data = county_covid_t, col_name = "NewCases")
total_deaths <- clean_column(data = county_covid_t, col_name = "TotalDeaths")
new_deaths <- clean_column(data = county_covid_t, col_name = "NewDeaths")
```

**Create a tibble from the TotalCases and TotalDeaths called cleaned_total_cases_deaths**
**Add a calculated column PercentDeaths that shows the death rate**
```{r}
cleaned_total_cases_deaths <- 
  tibble(TotalCases = total_cases,  
         TotalDeaths = total_deaths) %>%
  mutate(PercentDeaths = 
           paste0(round((total_deaths / total_cases) * 100, 1), "%"))
```

**Add county data to the cleaned_total_cases_deaths tibble and sort by county**
```{r}
cleaned_total_cases_deaths <- 
  bind_cols(county_covid_t %>% select(County), cleaned_total_cases_deaths) %>%
  arrange(County)
```


**Create a ggplot where the bars represent the total cases and the labels** 
**represent the death rate for each county**
**Use geom_text_repel for the death rate adn title the x labels 90 degrees to prevent overlap**
```{r}
ggplot(data = cleaned_total_cases_deaths, 
       mapping = aes(x = County, y = TotalCases)) +
  geom_bar(stat = "identity", fill = "magenta") +
  geom_text_repel(aes(label = PercentDeaths), max.overlaps = 25, size = 3.0) +
  theme(axis.text.x = element_text(angle = 90, hjust = 0.5)) +
  scale_x_discrete(expand = expansion(mult = 0.0)) +
  labs(title = "Countywise Total Cases and Percent Deaths for March 19, 2021")
```

**3.Identify the top two counties reporting most new cases.** 
```{r}
top_2_newcases_counties <-
  tibble(county_covid_t %>% select(County), new_cases) %>% 
  arrange(desc(new_cases)) %>%
  head(2)

top_2_newcases_counties  
```

# PROBLEM 2:
  
**Obtain your free API for https://spoonacular.com/food-api **
**Use it to obtain a list of 10 recipes that have carbohydrates not exceeding**
**30 grams. Present your output as a 10x3 tibble, where the column names are** 
**“Recipe” (the title of the recipe), “Carbs” (the carb content) and** 
**“ID” (the ID of the recipe)**
**Find 10 types of Riesling wines whose prices do not exceed $50 and present** 
**your results as a 10x3 tibble, where the columns** 
**represent the title of the wine, its ID and its price.**

**Define the API key to be used for both parts**
```{r}
api_key <-"87c3b12b27b04583b7fe359c409b8fd1"
```

**Construct the url to be used for the recipes**
```{r}
base_url_recipe <- "https://api.spoonacular.com/recipes/findByNutrients?"
query_recipe <- "maxCarbs=30&number=10&apiKey="
url_recipe <- paste0(base_url_recipe, query_recipe, api_key)
```

**Perform the API call, save the results into a tibble,extracting only the necessary**
**columns. Rename the columns to match the names in the problem statement**
```{r}
json_result_recipe <- url_recipe %>% 
  fromJSON() 

recipes_t <- json_result_recipe %>%
  as_tibble() %>%
  select(title, carbs, id) %>%
  rename(Recipe = title, Carbs = carbs, ID = id)

recipes_t
```

**Construct the url to be used for the recipes**
```{r}
base_url_wine <-"https://api.spoonacular.com/food/wine/recommendation?"
query_wine <- "wine=Riesling&maxPrice=50&number=10&apiKey="
url_wine <- paste0(base_url_wine, query_wine,api_key)
```

**Perform the API call**
```{r}
json_result_wine <- url_wine %>% 
  fromJSON() 
```

**Given the more complex output, extract the respective fields directly from the**
**the result of fromJSON(). Use successive $ to go deeper**
**Create a tibble from the fields requested in the problem statement**
```{r}
cleaned_wines_t <- tibble(Wine = json_result_wine$recommendedWines$title,
                          ID = json_result_wine$recommendedWines$id, 
                          Price = json_result_wine$recommendedWines$price)


cleaned_wines_t
```
