---
title: "Assignment 3 Data Wrangling"
author: "Yaniv Bronshtein"
date: "02/15/2021"
output: html_document
---
  
  ```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import the libraries
```{r}
library(tidyverse)
library(Lahman)
library(lubridate)
library(curl)
library(stringr)

```

## (1) We would like to create a data frame like the babynames data frame, 
## but for baseball players.



### (1a) Use the Master data frame in the Lahman package to create a tibble 
### with exactly the same variables as the babynames data frame, and ordered in \
### the same way. You will need to use the summarize() function to get the counts 
### of each name's use. For year, use the year of birth. For name, use the first 
### name (variable nameFirst). The final table should look like this 
### where prop is the proportion of names in a specific birthyear)

###------------------------------------------
  
###  birthYear   nameFirst    n  prop

###-------------------------------------------


**Call group by to group master by birthYear and nameFirst**
**Then, pipe the result to summarise() which will return a tibble with birthYear**
**nameFirst, and a column by the name n with the counts of each grouping**
**Pipe the result to mutate() and create a calculated column prop, rounding the**
**value to 3 digits after the decimal**
**Call arrange() to sort by the birthYear ascending**
```{r}
Master_t <- as_tibble(Master)
name_by_year <- Master_t %>%
  group_by(birthYear, nameFirst) %>%
  summarise(n = n()) %>%
  mutate(prop = round(n / sum(n), digits = 3)) %>%
  arrange(birthYear)
name_by_year
```



### (1b) In the Master dataframe, let us check whether the variable birthYear is 
### consistent with the year in birthDate. 
### Use a function in the lubridate package (discussed in Lecture 2) to extract 
### the year from the birthDate. Call this variable birthYear2. 
### In how many cases does birthYear have an "NA" entry? 
### In how many cases does birthYear2 have an "NA" entry? 
### In how many cases do both have "NA" entries? 
### if you ignore all the cases with at least one "NA" entry 
### (either in the birthYear or birthYear2 variable), 
### do all the remaining cases match?

**Create a temporary dataFrame**
```{r}
q1_t <- Master_t %>% select(playerID, birthYear, birthDate) %>%
  mutate(birthYear2 = year(birthDate))
```

**In how many cases does birthYear have an "NA" entry**
```{r}
count_na_birthYear <- q1_t %>% select(birthYear) %>% 
  is.na() %>% sum() %>% as.numeric()
count_na_birthYear
```
**In how many cases does birthYear2 have an "NA" entry?**
```{r}
count_na_birthYear2 <- q1_t %>% select(birthYear2) %>% 
  is.na() %>% sum() %>% as.numeric()
count_na_birthYear2
```
**In how many cases do both have "NA" Entries**
```{r}
count_both_na <- q1_t %>% 
  filter(is.na(birthYear), is.na(birthYear2)) %>% nrow() %>% as.numeric()
count_both_na
```
**If you ignore all cases with at least one "NA" entry,, do all the remaining**
**Cases match?**
**To answer this, first create a temporary data frame neither_na_t**
**Use it to compute count_neither_na and count_matching_after_neither_na**
```{r}
neither_na_t <- q1_t%>% 
  filter(!(is.na(birthYear) | is.na(birthYear2)))
count_neither_na <- neither_na_t %>% nrow() %>% as.numeric()
count_neither_na

count_matching_after_neither_na <- neither_na_t %>% 
  filter(as.numeric(birthYear) == as.numeric(birthYear2)) %>% nrow() %>% as.numeric()
count_matching_after_neither_na
```

*As you can see based on the results, the count is the same, meaning that all*
*the remaining cases match*


### (1c)  Create a data frame of players showing just the playerID, first name, 
### last name, given name, and career total 
### (meaning, summed over all years and all stints) of games 
### (that is, the G variable) according to the Fielding data frame.

**Create a tibble to store Fielding data**
```{r}
fielding_t <- as_tibble(Fielding)
```
**Create a variable FieldingSmall containing only playerID and G from fielding_t**
```{r}
FieldingSmall <- fielding_t %>% select(playerID, G)
```

**Create a variable Master_small_t containing only playerID, nameFirst, nameLast**
**and name given from Master_t**
```{r}
Master_small_t <- Master_t %>% select(playerID, nameFirst, nameLast, nameGiven)
```

**Perform a left join between the two subset tables on the key playerID**
**Then, pipe the result to group by playerID, nameFirst, nameLast, nameGiven**
**Pipe the result after grouping to create a final tibble with added total games**
**column**
```{r}
player_t <- FieldingSmall %>% left_join(Master_small_t, by = "playerID") %>%
  group_by(playerID, nameFirst, nameLast, nameGiven) %>% summarise(total_games = sum(G))
player_t
```

### (1d) Using mutate() and str_c(), add a variable to your data frame in (c) for full name by 
### combining the first name and last name with a space between them.
```{r}
player_t <- player_t %>% mutate(fullName = 
                                       str_c(nameFirst, nameLast, sep = " "))
player_t
```


### (1e) Use the data frames you've created to determine the 5 most popular
### first names in baseball among players who played at least 500 games. 
### Plot them over time with lines in a single plot. 
### Be sure to make the plot look nice by using a title and changing the 
### axis labels if necessary.


**Compute the name_count by filtering player_t to only players over 500 games,**
**piping the result to ungroup() in order to calculate the count of first names**
**piping the result to arrange by n in descending, and finally using slice to**
**split to top 5**
```{r}
name_count <- player_t %>% filter(total_games >= 500) %>%
  ungroup() %>%
  count(nameFirst) %>%
  arrange(desc(n)) %>%
  slice(1:5)
```
**Generate the plot**
```{r}
ggplot(data = name_count, 
       mapping = aes(x = nameFirst, y = n, group = 1)) +
  geom_line() +
  geom_point() +
  labs(
    title = "5 most popular first name with at least 500 games",
    x = "First Name",
    y = "Name Count"
  ) +
  theme(plot.title = element_text(hjust = 0.5))
```


**2) Read the post at http://www.sumsar.net/blog/2016/09/whats-on-the-menu/ and** 
**follow the steps yourself.**

**(Please include the R code in the RMarkdown file up through the creation of** 
**the data frame d - a terrible name, by the way.)**

**Would you have plotted any of the graphs on the webpage differently?**

**The author of the blog post found some interesting, or at least amusing,** 
**things in the data. Replacing the boring "tea",**
**"coffee", "apple" "banana", explore the rise or fall of more interesting**
**things like noodles, tandoori, curry, sushi and**
**kale in the menu.**
```{r}
menu_data_url <- 
  "https://s3.amazonaws.com/menusdata.nypl.org/gzips/2016_09_16_07_00_30_data.tgz"
temp_dir <- tempdir()
curl_download(menu_data_url, file.path(temp_dir, "menu_data.tgz"))
untar(file.path(temp_dir, "menu_data.tgz"), exdir = temp_dir)
dish <- read_csv(file.path(temp_dir, "Dish.csv"))
menu <- read_csv(file.path(temp_dir, "Menu.csv"))
menu_item <- read_csv(file.path(temp_dir, "MenuItem.csv"))
menu_page <- read_csv(file.path(temp_dir, "MenuPage.csv"))


d <- menu_item %>% select( id, menu_page_id, dish_id, price) %>%
  left_join(dish %>% select(id, name) %>% rename(dish_name = name),
            by = c("dish_id" = "id")) %>%
  left_join(menu_page %>% select(id, menu_id),
            by = c("menu_page_id" = "id")) %>%
  left_join(menu %>% select(id, date, place, location),
            by = c("menu_id" = "id")) %>%
  mutate(year = lubridate::year(date)) %>%
  filter(!is.na(year)) %>%
  filter(year > 1800 & year <= 2016) %>%
  select(year, location, menu_id, dish_name, price, place)




ggplot(d, aes(year)) +
  geom_histogram(binwidth = 5, center = 1902.5, color = "black", fill = "lightblue") +
  scale_y_continuous("N.o. menu items")


d$decennium = floor(d$year / 10) * 10
#foods <- c("coffee", "tea", "pancake", "ice cream", "french frie",
#           "french peas", "apple", "banana", "strawberry")

foods <- c("noodle", "tandoori", "curry", "sushi", "kale")

food_over_time <- map_df(foods, function(food) {
  d %>%
    filter(year >= 1900 & year <= 1980) %>%
    group_by(decennium, menu_id) %>%
    summarise(contains_food =
                any(str_detect(dish_name, regex(paste0("\\b", food), ignore_case = TRUE)),
                    na.rm = TRUE)) %>%
    summarise(prop_food = mean(contains_food, na.rm = TRUE)) %>%
    mutate(food = food)
})



# A reusable list of ggplot2 directives to produce a lineplot
food_time_plot <- list(
  geom_line(),
  geom_point(),
  scale_y_continuous("% of menus include",labels = scales::percent,
                     limits = c(0, NA)),
  scale_x_continuous(""),
  facet_wrap(~ food),
  theme_minimal(),
  theme(legend.position = "none"))
```

**Here is my own exploration of the various foods in the list noodles,**
**tandoori, curry, sushi, and kale**

```{r}
food_over_time %>% filter(food %in% c("noodle", "tandoori")) %>%
  ggplot(aes(decennium, prop_food, color = food)) + food_time_plot

```


```{r}
food_over_time %>% filter(food %in% c("noodle", "sushi")) %>%
  ggplot(aes(decennium, prop_food, color = food)) + food_time_plot

```


```{r}
food_over_time %>% filter(food %in% c("noodle", "kale")) %>%
  ggplot(aes(decennium, prop_food, color = food)) + food_time_plot

```

*I would have plotted the above graphs as different colored lines on the same*
*plot instead of each one on a different plot*
*This first plot demonstrates that tandoori only started getting popular in the*
*1980's*

*the second plot demonstrates that noodes have generally been popular compared*
*to sushi*

*In the last plot, it can be seen that Kale at it's peak comprised less than 2.5%*
*of the data*
