---
title: "Data Wrangling Assignment 5"
author: "Yaniv Bronshtein"
date: "2/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Import the necessary libraries**  
```{r}
library(tidyverse)
library(broom)
library(gapminder)
library(rsample)
```

# Problem 1
In this exercise we will work with the total number of words spoken by characters of different races and genders in the Lord of the Rings movies.  
1.Get the data in a single data frame  
Create 3 data frames (or tibbles) from these files:  
```{r}
fellowship_t <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Fellowship_Of_The_Ring.csv") %>% as_tibble()

two_towers_t <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Two_Towers.csv") %>% as_tibble()

return_t <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/The_Return_Of_The_King.csv") %>% as_tibble()
```

2. Tidy the combined data frame by creating new variables “Gender” and “Words”  
**Use bind_rows to sequentially append all the tibbles to each other**
**Then, use pivot_longer() to convert the Male and Female columns to values in the Gender column**
**In turn, move the values originally in these columns to a new column called Words**
```{r}
lotr_t <- fellowship_t%>%bind_rows(two_towers_t) %>% bind_rows(return_t)
lotr_t <- lotr_t %>% pivot_longer(cols = c("Female", "Male"), names_to = "Gender", values_to = "Words")
```

3.Use the combined data frame to answer the following questions  
a)How many words were spoken in each movie?  
**First, create a tibble by grouping lotr_t by Film and then using summarise() to create a new column total_words based**
**on the film groups**
```{r}
total_movie_wc_t <- lotr_t %>% group_by(Film) %>% summarise(total_words=sum(Words))
```
**Next, derive the count in the first film by using str_detect on the Film column to filter the dataframe,**
**pulling out the column in vector form, and then converting the result to a numeric value**
**Repeat thuis step for the other two movies**
```{r}
count_fellowship <- total_movie_wc_t %>% filter(str_detect(Film, "Fellow")) %>% pull(total_words) %>% as.numeric()
cat("The Fellowship of the Ring Total Word Count", count_fellowship)

count_towers <- total_movie_wc_t %>% filter(str_detect(Film, "Tower")) %>% pull(total_words) %>% as.numeric()
cat("The Two Towers Total Word Count", count_towers)

count_return <- total_movie_wc_t %>% filter(str_detect(Film, "Return")) %>% pull(total_words) %>% as.numeric()
cat("The Return of the King Total Word Count", count_return)
```

b)How many words were spoken by each gender in total?  
**First, create a tibble by grouping lotr_t by Gender and then using summarise() to create a new column words_by_gender based on the Gender groups**
```{r}
words_by_gender_t <- lotr_t %>% group_by(Gender) %>% summarise(words_by_gender = sum(Words))
```

**Next, derive the count of males and females by filtering on Male and Female respectively**
**pulling out the column in vector form, and then converting the result to a numeric value**
```{r}
male_wc <- words_by_gender_t %>% filter(Gender=="Male") %>% pull(words_by_gender) %>% as.numeric()
cat("Total words spoken by men", male_wc)

female_wc <- words_by_gender_t %>% filter(Gender=="Female") %>% pull(words_by_gender) %>% as.numeric()
cat("Total words spoken by women", female_wc)
```

c)How many words were spoken by each race in total?  
**First, create a tibble by grouping lotr_t by Race and then using summarise() to create a new column words_by_race based on the Race groups**
```{r}
words_by_race_t <- lotr_t %>% group_by(Race) %>% summarise(words_by_race = sum(Words))
```

**Next, derive the count of each race filtering the Race column,**
**pulling out the column in vector form, and then converting the result to a numeric value**
```{r}
elf_wc <- words_by_race_t %>% filter(Race == "Elf") %>% pull(words_by_race) %>% as.numeric()
cat("Total words spoken by elves", elf_wc)

hobbit_wc <- words_by_race_t %>% filter(Race == "Hobbit") %>% pull(words_by_race) %>% as.numeric()
cat("Total words spoken by Hobbits", hobbit_wc)

man_wc <- words_by_race_t %>% filter(Race == "Man") %>% pull(words_by_race) %>% as.numeric()
cat("Total words spoken by Man", man_wc)
```

4.Create a data frame with totals by race and movie, calling it by_race_film.
```{r}
by_race_film_t <- lotr_t %>% group_by(Film, Race) %>% summarise(words_by_race_movie = sum(Words))
by_race_film_t
```

# Problem 2
1.Split/group the gapminder data by country. For each country,   
fit an ARIMA(0,0,1) or MA(1) model to lifeExp, and produce a tibble that   
the country-wise values of AIC and BIC, two measures of goodness of   
model fit. Obtain a scatter plot of AIC versus BIC and comment.  


**Convert gapminder data set to tibble form**
```{r}
gapminder_t <- gapminder %>% as_tibble()
```

**Create a function compute_aic_bic() to generate a tibble containing the country name,**
**AIC, and BIC in that order**
```{r}
#This function returns a tibble containing the country, AIC, BIC
# Params:
#p1: number of AR coefficients 
#p2: number of differences
#p3: number of MA coefficients
compute_aic_bic = function(p1, p2, p3) {
  
  #Split the gapminder_t by country using group_split().
  #Use map to apply arima() with order based on function input
  #Use map again to apply the broom function glance() to receive model level information
  #(AIC and BIC)
  countries_arima <- gapminder_t %>% 
  group_by(country) %>%
  group_split() %>%
  map(~arima(.$lifeExp, order = c(p1, p2, p3))) %>%
  map(glance)

  #Extract AIC and BIC in vector form using map_dbl()
  countries_aic <- countries_arima %>% map_dbl(~.$AIC)
  countries_bic <- countries_arima %>% map_dbl(~.$BIC)
  #Create a tibble by combining the countries(removing duplicates using the unique())
  #function, and adding the extracted AIC and BIC to the end
  countries_t <- gapminder_t %>% select(country) %>% 
  unique() %>% 
  mutate(AIC = countries_aic) %>%
  mutate(BIC = countries_bic)
  
  return(countries_t)
}
```

**Generate the ggplot of AIC vs BIC for ARIMA(0,0,1)**
```{r}
m1 <- compute_aic_bic(0, 0, 1)
ggplot(data = m1 , mapping = aes(x = AIC, y = BIC)) +
  geom_point(color = "Magenta") +
  labs(
    title = "BIC vs AIC for countries in Gapminder dataset"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))

```

2.Now repeat the previous step for four other models: ARIMA(0,0,1), ARIMA(0,0,2),   
ARIMA(0,0,3), ARIMA(0,1,0), ARIMA(0,1,1), and in a single plot, show boxplots   
of AIC values for the five models. Based on the boxplot, which of these five   
models do you think fits the data best for most countries?  

**In the code below, we will call compute_aic_bic() for the required models**
**And create an additional column filled with the corresponding string**
**to make plotting easier**
```{r}
m1 <- m1 %>% mutate(Model = "ARIMA(0,0,1)" )
m2 <- compute_aic_bic(0, 0, 2) %>% select(AIC) %>% mutate(Model = "ARIMA(0,0,2)") 
m3 <- compute_aic_bic(0, 0, 3) %>% select(AIC) %>% mutate(Model = "ARIMA(0,0,3)")
m4 <- compute_aic_bic(0, 1, 0) %>% select(AIC) %>% mutate(Model = "ARIMA(0,1,0)")
m5 <- compute_aic_bic(0, 1, 1) %>% select(AIC) %>% mutate(Model = "ARIMA(0,1,1)")
```

**Next, we stack each model on top of one another to create a unified tibble**
**with the following structure:**
**Country|AIC|BIC|Model**
```{r}
models_t <- bind_rows(m1, m2, m3, m4, m5) 
```

**Generate the boxplot**
```{r}
ggplot(data = models_t, mapping = aes(x = Model, y = AIC)) + 
  geom_boxplot() + 
  labs(title = "AIC distribution by Arima Model")
```
*The best model has the minimal AIC, so ARIMA(0,1,1) wins*

3.Filter the data only for continent Europe.   
For the best model identified in step 2, create a tibble showing the   
country-wise model parameters (moving average coefficients) and their   
errors using the broom package.  

**Create a separate dataframe on gapminder just for Europe**
```{r}
gapminder_europe <- gapminder_t %>% filter(continent == "Europe")
```
**Apply similar logic to that of compute_aic_bic(), replacing glance() with tidy()**
**to generate coefficient level data**

```{r}
countries_arima_europe_tidy <- gapminder_europe %>%
  group_by(country) %>%
  group_split() %>%
  map(~arima(.$lifeExp, order = c(0, 1, 1))) %>%
  map(tidy)

countries_ma_estimate <- countries_arima_europe_tidy %>% 
  map_dbl(~.$estimate)
countries_ma_std.error <- countries_arima_europe_tidy %>% 
  map_dbl(~.$std.error)

countries_list <- gapminder_europe %>% select(country) %>% unique()
countries_ma_estimate_error_t <- 
  tibble(countries_list, 
         estimate = countries_ma_estimate, error = countries_ma_std.error)

countries_ma_estimate_error_t
```

4.Now filter the data only for year 1992. Plot lifeExp against log10(gdpPercapita). 
Fit a linear model of lifeExp on log10(gdpPercapita) using population as weights 
and obtain (i) bootstrapped 95% confidence intervals for the slope coefficient 
and (ii) bootstrapped 90% prediction intervals for each data point using 500 
bootstrapped samples (show a plot of the prediction intervals). 
Compare the bootstrapped 95% confidence intervals for the estimated slope 
coefficient with those generated automatically by the lm() function. 
Which one is wider?

**Step 1: create a filtered tibble on gapminder_t for the year 1992**
```{r}
gapminder_1992_t <- gapminder_t %>% filter(year == 1992)
```

**Step 2: Generate a linear model to be used as reference and display both coefficient and confidence interval data**
```{r}
lm_1992 <- gapminder_1992_t %>% 
  lm(lifeExp ~ log10(gdpPercap), weights = pop, data = .)

lm_1992_coef <- coef(lm_1992)
cat("LM 1992 coefficients:\n", lm_1992_coef,"\n")
lm_1992_confint <- confint(lm_1992, level = 0.95)
cat("LM 1992 confidence interval:\n", lm_1992_confint,"\n")
```

**Step 3: Generate a scatter plot showing Life Expectancy vs log10(gdpPercapita)**
```{r}
ggplot(data = gapminder_1992_t, 
       mapping = aes(x = log10(gdpPercap), y = lifeExp)) +
  geom_point() + 
  labs(
    title = "Life Expectancy vs log10(gdpPercapita) for gapminder 1992 Data"
  ) + 
  theme(plot.title = element_text(hjust = 0.5))
```

**Step 4: As part of the requirement for part (i) provide bootstrapped 95% confidence intervals for the slope coefficient. Tidy is necessary for this**
```{r}
set.seed(1)
alpha1 = 0.05
boot_lm <- gapminder_1992_t %>% bootstraps(500)
boot_lm1 <- map(boot_lm$splits, ~as_tibble(.)) %>%
  map(~tidy(lm(lifeExp ~ log10(gdpPercap), weights = pop, data = .))) %>%
  bind_rows(.)
```
**Step 4a. Display the confidence interval together with the median for part (i)**
```{r}
conf_int_95 <- boot_lm1 %>% 
  group_by(term) %>% 
  summarise(conf.low = quantile(estimate, alpha1 /2),
                                      conf.high = quantile(estimate, 1 - alpha1 /2),
                                      median = median(estimate))
conf_int_95
```

**Step 5: As part of the requirement for part (ii), provide the bootstrapped 90% prediction intervals**
**Augment() is now necessary for this**
```{r}
alpha2 = 0.1
boot_lm2 <- map(boot_lm$splits, ~as_tibble(.)) %>%
  map(~augment(lm(lifeExp ~ log10(gdpPercap), weights = pop, data = .))) %>%
  # Needed to rename because of special characters
  bind_rows(.) %>% rename(log_gdp_percap = names(.)[2]) 
```

**Step 5a. Display the confidence interval together with the median for part (ii)**
```{r}
conf_int_90 <- boot_lm2 %>% group_by(log_gdp_percap) %>%
  summarise(conf.low = quantile(.fitted, alpha2 / 2), 
            conf.high = quantile(.fitted, 1 - alpha2 / 2), 
            median = median(.fitted))

conf_int_90
```

**Show a plot of the prediction intervals**
```{r}
ggplot(conf_int_90) + 
  geom_point(aes(x = log_gdp_percap, y = lifeExp), data = boot_lm2) +
  geom_line(aes(x = log_gdp_percap, y = median)) +
  geom_ribbon(aes(x = log_gdp_percap, ymin = conf.low, ymax = conf.high), 
              alpha = 0.1) +
  labs(
    title = "Plot of prediction intervals for bootstraps(500) of a Linear Model",
    x = "log10(gdpPercap)")
```
*I believe that the bootstrapped is wider*